---
title: "Thesis Modeling"
author: "Tate Huffman"
date: "12/05/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(optimr)
library(nloptr)
library(knitr)
library(kableExtra)
bat = read_csv('../Data/Clean/bat/bat_all.csv')
pitch = read_csv('../Data/Clean/pitch/pitch_all.csv')
```

## Model Setup

This first regression is minimizing the objective function

$$Q = \sum_{i=1}^{20} [P_i - \pi_i(\Omega)]^2$$

with respect to the parameters of $\Omega$. The values of $i$ correspond to the exit or promotion of the individual in a given time period: for $1 \leq i \leq 10$, this denotes an exit in year $i$, and for $11 \leq i \leq 20$, this denotes a promotion of this person in year $i-10$ (i.e., $i=12$ means a promotion for the individual in year 2). $P_i$ is the observed probability of a worker exiting or promoting in period $i$.

$\pi_i(\Omega)$ is the predicted probability of an exit or promotion in period $i$, where $\Omega = (\alpha, \tau, \theta_0, \mu, \sigma)$. These variables, respectively, correspond to: the probability of a Type A (able) worker producing a good signal; the probability of a Type B (unable) worker producing a good signal; the ex ante probability of a new worker being Type A; the mean of the worker promotion threshold $\theta_u$; and its standard deviation.

When simplified, we see the probability of an exit in period $i$ as

$$\pi_i = [\theta_0 \alpha^{i-1}(1-\alpha) + (1-\theta_0) \tau^{i-1}(1-\tau)] \int_{\theta (i-1,i-1)}^{1} g(\theta_u; \mu, \sigma) d\theta_u$$
and the probability of a promotion in period $i$ as

$$\pi_{i+10} = [\theta_0 \alpha^i + (1-\theta_0) \tau^i] \int_{\theta (i-1,i-1)}^{\theta (i,i)} g(\theta_u; \mu, \sigma) d\theta_u$$
where $g(\theta_u; \mu, \sigma)$ is the beta distribution of $\theta_u$. Additionally, $\theta(x,n)$ represents the probability of a worker producing $x$ good signals in $n$ periods, which by Bayes' rule can be represented as

$$\theta(x,n) = \frac{\theta_0 \alpha^x (1-\alpha)^{n-x}}{\theta_0 \alpha^x (1-\alpha)^{n-x} + (1-\theta_0) \tau^x (1-\tau)^{n-x}}$$

## Initial Regression

For this first attempt at the regression, we use its simplest form. This means that the statistics we use to qualify as a good signal are not normalized to league, and in fact are generalized across leagues: a good hitting signal is defined as an OPS over .825, and good hitting signal is defined as an ERA under 3.75. In future regressions, this performance will have appearance minimums, use more representative statistics, and normalize performance to the league's level of offense, but for now, this is a very rough first pass.

```{r add_vars, echo=FALSE, message=FALSE}

# Number of years and performance benchmark for "good signal"
# OPS of 0.700, ERA of 4.50 (these are loose assumptions)

n_years <- 10
bat_std <- 0.700
pitch_std <- 4.50

# Batters

bat_performance <- bat %>% 
  # making level into scalar for ease of comparison
  mutate(Level = case_when(Level == "Rookie" ~ 1,
                           Level == "Short-Season A" ~ 2,
                           Level == "A" ~ 3,
                           Level == "Adv A" ~ 4,
                           Level == "AA" ~ 5,
                           Level == "AAA" ~ 6,
                           Level == "MLB" ~ 7),
         # good signal is when OPS is greater than standard
         signal = if_else(OPS >= bat_std, 1, 0, missing = 0)) %>% 
  group_by(Name) %>% 
  mutate(yr_unique = cumsum(!duplicated(Year)), # counter of unique year played
         exit = if_else(yr_unique == max(yr_unique), 1, 0)) %>% 
  ungroup() %>% 
  group_by(Name, Year) %>% 
  # promotion if they either played at multiple levels or played at MLB
  # latter is because there's no higher stage than MLB...
  # ... so staying at this level is equivalent of constant promotion
  # ** potentially faulty assumption, TBD **
  # mutate(promotion = if_else(n_distinct(Level) != 1 | max(Level) == 7, 1, 0)) %>%
  mutate(promotion = if_else(n_distinct(Level) != 1, 1, 0)) %>%
  ungroup() %>%
  select(Name, Age, Year, Level, signal, yr_unique, exit, promotion) %>% 
  # limiting to ten seasons of data, another assumption
  filter(yr_unique <= n_years) %>%
  group_by(Name, Year) %>% 
  summarize(Age = max(Age),
            Level = max(Level),
            signal = max(signal),
            yr_unique = max(yr_unique),
            exit = max(exit),
            promotion = max(promotion)) %>% 
  group_by(Name) %>% 
  # also says someone was promoted if their level was higher this year than before
  mutate(promotion = if_else(Level > lag(Level), 1, promotion),
         signal_ct = cumsum(signal)) %>% 
  ungroup() %>% 
  replace_na(list(promotion = 0))

# Pitchers

pitch_performance <- pitch %>% 
  # making level into scalar for ease of comparison
  mutate(Level = case_when(Level == "Rookie" ~ 1,
                           Level == "Short-Season A" ~ 2,
                           Level == "A" ~ 3,
                           Level == "Adv A" ~ 4,
                           Level == "AA" ~ 5,
                           Level == "AAA" ~ 6,
                           Level == "MLB" ~ 7),
         # good signal is when OPS is greater than standard
         signal = if_else(ERA <= pitch_std, 1, 0, missing = 0)) %>% 
  group_by(Name) %>% 
  mutate(yr_unique = cumsum(!duplicated(Year)), # counter of unique year played
         exit = if_else(yr_unique == max(yr_unique), 1, 0)) %>% 
  ungroup() %>% 
  group_by(Name, Year) %>% 
  # promotion if they either played at multiple levels or played at MLB
  # latter is because there's no higher stage than MLB...
  # ... so staying at this level is equivalent of constant promotion
  # ** potentially faulty assumption, TBD **
  # mutate(promotion = if_else(n_distinct(Level) != 1 | max(Level) == 7, 1, 0)) %>%
  mutate(promotion = if_else(n_distinct(Level) != 1, 1, 0)) %>%
  ungroup() %>%
  select(Name, Age, Year, Level, signal, yr_unique, exit, promotion) %>% 
  # limiting to ten seasons of data, another assumption
  filter(yr_unique <= n_years) %>%
  group_by(Name, Year) %>% 
  summarize(Age = max(Age),
            Level = max(Level),
            signal = max(signal),
            yr_unique = max(yr_unique),
            exit = max(exit),
            promotion = max(promotion)) %>% 
  group_by(Name) %>% 
  # also says someone was promoted if their level was higher this year than before
  mutate(promotion = if_else(Level > lag(Level), 1, promotion),
         signal_ct = cumsum(signal)) %>% 
  ungroup() %>% 
  replace_na(list(promotion = 0))

```

-- RIGHT NOW: Future modification: instead of doing P(good signal) = x, P(bad signal) = 1-x, reverse it so that P(bad signal) = y, P(good signal) = 1-y -> professional baseball is more of a war of attrition, so if you're always not bad then that will get you promoted as people retire... use player IDs instead of names --

-- FOR PAPER: what standard to use for OPS/ERA? should be relative to league... should the promotion variable be 1 if they remain in the major leagues? (it's not currently) don't we want to weight the major leagues heavier?... explain upper/lower bounds, and have system for more rigorous initial guess... the promotion probability threshold is a mean across all levels, not specified level-by-level... intuition behind very accurate exit probabilities, but inaccurate promotion probabilities - in all empirical analyses, the exit is a one-time thing, but not the case here for promotion -> skewing the data?... weighting promotion count by total number of promotions here to normalize within the data - is that valid? should we also be normalizing for promotion probabilities?... how does fit of the model change over time/with different stats? able to reveal changes in team tendencies over time?... updated tables for summary statistics and inclusion of hazard rate and transition matrix... exit should NOT just be the last year the player is in the data - should not include 2019 -- 

```{r exit_and_promotion, echo=FALSE, message=FALSE}

# Batter exits

n_bat <- bat %>% # number of unique batters
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

exit_distr_bat <- bat %>% # distribution of player exit times
  group_by(Name) %>%
  summarise(n_year = n_distinct(Year)) %>% 
  group_by(n_year) %>% 
  summarise(n_players = n_distinct(Name)) %>% 
  filter(n_year <= n_years) %>% 
  mutate(pct = n_players / n_bat,
         pos = 'Batters') # 100x multiplier to make it a true percent

# Pitcher exits

n_pitch <- pitch %>% # number of unique pitchers
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

exit_distr_pitch <- pitch %>% # distribution of player exit times
  group_by(Name) %>%
  summarise(n_year = n_distinct(Year)) %>% 
  group_by(n_year) %>% 
  summarise(n_players = n_distinct(Name)) %>% 
  filter(n_year <= n_years) %>% 
  mutate(pct = n_players / n_pitch,
         pos = 'Pitchers') # 100x multiplier to make it a true percent

# Merging the dataframes

n_total <- n_bat + n_pitch

exit_distr <- exit_distr_bat %>% 
  bind_cols(exit_distr_pitch) %>% 
  mutate(n_year = n_year...1,
         n_players = n_players...2 + n_players...6,
         pct = n_players / n_total) %>% 
  select(n_year, n_players, pct)

# Batter promotions

promote_distr_bat <- bat_performance %>% 
  group_by(yr_unique) %>% 
  summarize(n_players = sum(promotion),
            pct = n_players / n_bat)

# Pitcher promotions

promote_distr_pitch <- pitch_performance %>% 
  group_by(yr_unique) %>% 
  summarize(n_players = sum(promotion),
            pct = n_players / n_pitch)

# Weighting the percent promoted by the average number of promotions
# This is to prevent probabilities from surpassing 1

performance <- bat_performance %>% 
  bind_rows(pitch_performance)

n_promotions <- performance %>% 
  group_by(Name) %>% 
  summarize(promotions_total = sum(promotion)) %>% 
  select(promotions_total) %>% 
  pull() %>% 
  mean()

# Combined promotions

promote_distr <- bat_performance %>% 
  group_by(yr_unique) %>% 
  summarize(n_players = sum(promotion)) %>% 
  bind_cols(pitch_performance %>% 
              group_by(yr_unique) %>% 
              summarize(n_players = sum(promotion))) %>% 
  mutate(n_years = yr_unique...1,
         n_players = n_players...2 + n_players...4,
         pct = (n_players / n_total) / n_promotions) %>% # weighting by total promotes
         # pct = (n_players / n_total)) %>% 
  select(n_years, n_players, pct)

# Defining the dataframe with probabilities for each cell

player_mvmt <- exit_distr %>% 
  select(pct) %>% 
  bind_rows(promote_distr %>% select(pct)) %>% 
  mutate(# pct = pct / sum(.$pct), # normalizing to sum to 1 - probably not accurate?
         year = rep(seq(1:10), 2),
         exit = c(rep(T, 10), rep(F, 10)),
         cell = seq(1:20))

```

```{r calc_functions, echo=FALSE, message=FALSE}

# want to define the function π_i(Ω) that can calculate it for each value of i

# Calculating bounds for the beta density integral

theta_bounds <- function(n, theta_0, alpha, tau){
  
  # n <- theta_inputs[1]
  # theta_0 <- theta_inputs[2]
  # alpha <- theta_inputs[3]
  # tau <- theta_inputs[4]

  num <- theta_0*(alpha^n)
  den <- theta_0*(alpha^n) + (1-theta_0)*(tau^n)
    
  return(num / den)
}

# Calculating the beta density integral
# Bounds change depending on whether it's an entry or exit

beta_integral <- function(exit, n, a, b, theta_0, alpha, tau){
  
  # exit <- beta_inputs[1]
  # n <- beta_inputs[2]
  # a <- beta_inputs[3]
  # b <- beta_inputs[4]
  # theta_0 <- beta_inputs[5]
  # alpha <- beta_inputs[6]
  # tau <- beta_inputs[7]
  
  bound_lower <- theta_bounds(n-1, theta_0, alpha, tau)
  bound_upper <- if_else(exit, 1, theta_bounds(n, theta_0, alpha, tau))
  
  integrand <- function(x){
    
    num <- (x^(a-1)) * ((1-x)^(b-1))
    den <- beta(a, b)
    
    return(num / den)
  }
  
  return(integrate(integrand, lower = bound_lower, upper = bound_upper)$value)
}

# Gets the multiplication term

mult_term <- function(exit, n, theta_0, alpha, tau){
  
  # exit <- mult_inputs[1]
  # n <- mult_inputs[2]
  # theta_0 <- mult_inputs[3]
  # alpha <- mult_inputs[4]
  # tau <- mult_inputs[5]
  
  val <- if_else(exit, 
                 theta_0*(alpha^(n-1))*(1-alpha) + (1-theta_0)*(tau^(n-1))*(1-tau),
                 theta_0*(alpha^n) + (1-theta_0)*(tau^n))
  
  return(val)
}

```

```{r optimize, echo=FALSE, message=FALSE}

# Function to optimize

fn_optim <- function(input_vals){ # making this a vector for optimization
  
  theta_0 <- input_vals[1]
  alpha <- input_vals[2]
  tau <- input_vals[3]
  a <- input_vals[4]
  b <- input_vals[5]
  
  fn_val <- 0
  
  for(i in 1:(2*n_years)){
    
    # Filtering the observed data to the cell in question
    
    data <- player_mvmt %>% 
      filter(cell == i)
    
    # Isolating variables for ease of use
    
    obsv <- data$pct # observed probability
    cell <- data$cell # cell index
    exit <- if_else(cell <= n_years, T, F) # whether it's an exit period
    
    # Value for predicted probability
    
    add_val <- mult_term(exit, i, theta_0, alpha, tau) * 
      beta_integral(exit, i, a, b, theta_0, alpha, tau)
    
    fn_val <- fn_val + (obsv - add_val)^2
  }
  
  return(fn_val)
}

# Inequality function

fn_ineq <- function(input_vals){
  
  return(input_vals[3] - input_vals[2]) # ensures that alpha >= tau
  
}

# Initial guesses for parameter values and their respective bounds
# theta_0, alpha, tau, a, b

init_guess <- c(0.05, 0.5, 0.5, 2, 2)
param_bounds_low <- c(0.01, 0.05, 0.05, 1, 1)
param_bounds_high <- c(1, 0.95, 0.95, Inf, Inf)

# Getting some results!

# Unconstrained optimization

# results <- optim(init_guess, fn_optim,
#                  lower = param_bounds_low,
#                  upper = param_bounds_high,
#                  method = "L-BFGS-B")

# Constrained optimization

results <- nloptr(init_guess,
                  fn_optim,
                  lb = param_bounds_low,
                  ub = param_bounds_high,
                  eval_g_ineq = fn_ineq,
                  opts = list("algorithm" = "NLOPT_LN_COBYLA",
                              "xtol_rel" = 1.0e-8))

# Getting the parameters from our calculations

params <- results$solution
theta_0_pred <- params[1]
alpha_pred <- params[2]
tau_pred <- params[3]
a_pred <- params[4]
b_pred <- params[5]

```

```{r figures, echo=FALSE}

# Adding predicted percent of players falling into cell

player_mvmt <- player_mvmt %>% 
  rowwise() %>% # applying the model row-by-row
  mutate(pred_pct = mult_term(exit, year, theta_0_pred, alpha_pred, tau_pred) * 
           beta_integral(exit, year, a_pred, b_pred, 
                         theta_0_pred, alpha_pred, tau_pred)) %>% 
  ungroup() %>% 
  pivot_longer(cols = ends_with("pct"),
               names_to = "observed",
               values_to = "pct") %>% 
  mutate(observed = if_else(str_detect(observed, "pred"), F, T))

# Graphing predicted percentages and observed percentages

# Exit probability

player_mvmt %>% 
  filter(exit) %>% 
  ggplot() +
  aes(x = year, y = 100*pct, color = observed) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,10,2)) +
  scale_y_continuous(breaks = seq(0,30,5)) +
  scale_color_manual(name = "Probability",
                     labels = c("Predicted", "Observed"),
                     values = c("red", "blue")) +
  labs(x = "Number of Seasons Played",
       y = "Exit Probability (Percent)",
       caption = "Levels: Rookie through MLB") +
  theme_classic()

# Promotion probability

player_mvmt %>% 
  filter(!exit) %>% 
  ggplot() +
  aes(x = year, y = 100*pct, color = observed) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,10,2)) +
  scale_y_continuous(breaks = seq(0,30,5)) +
  scale_color_manual(name = "Probability",
                     labels = c("Predicted", "Observed"),
                     values = c("red", "blue")) +
  labs(x = "Number of Seasons Played",
       y = "Promotion Probability (Percent)",
       caption = "Levels: Rookie through MLB") +
  theme_classic()

# Comparing observed exit and promotion probabilities

player_mvmt %>% 
  filter(observed) %>% 
  ggplot() +
  aes(x = year, y = 100*pct, color = exit) +
  geom_line() +
  scale_x_continuous(breaks = seq(0,10,2)) +
  scale_y_continuous(breaks = seq(0,30,5)) +
  scale_color_manual(name = "Probability",
                     labels = c("Promotion", "Exit"),
                     values = c("red", "blue")) +
  labs(x = "Number of Seasons Played",
       y = "Probability (Percent)",
       caption = "Levels: Rookie through MLB") +
  theme_classic()

```

```{r summary_bat, echo=FALSE}

# Updated tables from D3

# Compiling some batter summary statistics
# First getting a table w/ all years w/ >= 20 PA

bat_min_pa_20 <- bat %>% 
  group_by(Name, Year) %>% 
  mutate(pa = sum(PA)) %>% 
  filter(pa >= 20) %>% 
  ungroup()

# Number of players, both with minimums and without

n_bat <- bat %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

n_bat_min <- bat_min_pa_20 %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

# Number of years played, both with minimums and without

n_years <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

n_years_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

# Number of teams played for, both with minimums and without

n_teams <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

n_teams_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

# Number of levels played at, both with minimums and without

n_levels <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

n_levels_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

# Number of plate appearances per year, both with minimums and without

n_pa <- bat %>% 
  group_by(Name, Year) %>% 
  summarize(pa = sum(PA)) %>% 
  select(pa) %>% 
  pull()

n_pa_min <- bat_min_pa_20 %>% 
  group_by(Name, Year) %>% 
  summarize(pa = sum(PA)) %>% 
  select(pa) %>% 
  pull()

bat_player = tibble(
  Measure = c("Years Played", "Organizations", "Levels", "PA per Year"),
  M_1  = c(mean(n_years), mean(n_teams), mean(n_levels), mean(n_pa)),
  Pct1_10 = c(quantile(n_years, .1), quantile(n_teams, .1),
              quantile(n_levels, .1), quantile(n_pa, .1)),
  Pct1_90 = c(quantile(n_years, .9), quantile(n_teams, .9),
              quantile(n_levels, .9), quantile(n_pa, .9)),
  SD_1 = c(sd(n_years), sd(n_teams), sd(n_levels), sd(n_pa)),
  M_2  = c(mean(n_years_min), mean(n_teams_min), mean(n_levels_min), mean(n_pa_min)),
  Pct2_10 = c(quantile(n_years_min, .1), quantile(n_teams_min, .1),
              quantile(n_levels_min, .1), quantile(n_pa_min, .1)),
  Pct2_90 = c(quantile(n_years_min, .9), quantile(n_teams_min, .9),
              quantile(n_levels_min, .9), quantile(n_pa_min, .9)),
  SD_2 = c(sd(n_years_min), sd(n_teams_min), sd(n_levels_min), sd(n_pa_min))
)

```

```{r summary_pitch, echo=FALSE}

# Updated tables from D3

# Compiling some pitcher summary statistics
# First getting a table w/ all years w/ >= 20 BF

pitch_min_bf_20 <- pitch %>% 
  group_by(Name, Year) %>% 
  mutate(bf = sum(BF)) %>% 
  filter(bf >= 20) %>% 
  ungroup()

# Number of players, both with minimums and without

n_pitch <- pitch %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

n_pitch_min <- pitch_min_bf_20 %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

# Number of years played, both with minimums and without

n_years <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

n_years_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

# Number of teams played for, both with minimums and without

n_teams <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

n_teams_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

# Number of levels played at, both with minimums and without

n_levels <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

n_levels_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

# Number of batters faced per year, both with minimums and without

n_bf <- pitch %>% 
  group_by(Name, Year) %>% 
  summarize(bf = sum(BF)) %>% 
  select(bf) %>% 
  pull()

n_bf_min <- pitch_min_bf_20 %>% 
  group_by(Name, Year) %>% 
  summarize(bf = sum(BF)) %>% 
  select(bf) %>% 
  pull()

pitch_player = tibble(
  Measure = c("Years Played", "Organizations", "Levels", "BF per Year"),
  M_1  = c(mean(n_years), mean(n_teams), mean(n_levels), mean(n_bf, na.rm = TRUE)),
  Pct1_10 = c(quantile(n_years, .1), quantile(n_teams, .1),
              quantile(n_levels, .1), quantile(n_bf, .1, na.rm = TRUE)),
  Pct1_90 = c(quantile(n_years, .9), quantile(n_teams, .9),
              quantile(n_levels, .9), quantile(n_bf, .9, na.rm = TRUE)),
  SD_1 = c(sd(n_years), sd(n_teams), sd(n_levels), sd(n_bf, na.rm = TRUE)),
  M_2  = c(mean(n_years_min), mean(n_teams_min), mean(n_levels_min), mean(n_bf_min)),
  Pct2_10 = c(quantile(n_years_min, .1), quantile(n_teams_min, .1),
              quantile(n_levels_min, .1), quantile(n_bf_min, .1)),
  Pct2_90 = c(quantile(n_years_min, .9), quantile(n_teams_min, .9),
              quantile(n_levels_min, .9), quantile(n_bf_min, .9)),
  SD_2 = c(sd(n_years_min), sd(n_teams_min), sd(n_levels_min), sd(n_bf_min))
)

```

```{r transition_matrix, echo=FALSE}

total_matrix <- performance %>% 
  group_by(Name) %>% 
  filter(min(Year) %in% c(2008:2010)) %>% # only incl. those w/ full ten seasons
  mutate(start_level = first(Level)) %>% 
  ungroup() %>%
  group_by(start_level, yr_unique, Level) %>%
  summarize(ct = n_distinct(Name)) %>% 
  ungroup()

n_start <- total_matrix %>% 
  filter(yr_unique == 1) %>% 
  group_by(start_level) %>% 
  summarize(n = sum(ct))

total_matrix %>% 
  filter(yr_unique == 2, start_level == 7) %>% 
  add_row(start_level = 7,
          yr_unique = 2,
          Level = 0,
          ct = filter(n_start, start_level == 7)$n - sum(.$ct))

```

