---
title: 'Deliverable #3: Initial Data Exploration'
author: "Tate Huffman"
date: "10/29/2021"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
bat = read_csv('../Data/Clean/bat/bat_all.csv')
pitch = read_csv('../Data/Clean/pitch/pitch_all.csv')
```

```{r arrange_data, include=FALSE}

# Commenting this block out for now because we've already run it and modified the data 

# # Sorting the data so it looks prettier/easier to work with
# 
# bat <- bat %>% 
#   arrange(Name, Age, Year, desc(PA), Level, Org) %>% 
#   distinct()
# 
# pitch <- pitch %>% 
#   arrange(Name, Age, Year, desc(BF), Level, Org) %>% 
#   distinct()
# 
# # Overwriting the existing data
# 
# write_csv(bat, '../Data/Clean/bat/bat_all.csv')
# write_csv(pitch, '../Data/Clean/pitch/pitch_all.csv')
```

## Overview

The data collected for this thesis was obtained from [Baseball Reference](https://www.baseball-reference.com), a website that contains historical data for minor and major league baseball. The data was scraped using Python, the code for which can be found in my [GitHub repository](https://github.com/tjhuffman4246/senior-thesis). It contains performance data in minor and major league baseball from 1998 through 2019 and is separated into batting and pitching data.

## Format 

The data was originally separated into thousands of different files by team, based on batting/pitching, year, level of baseball (e.g., Triple-A, MLB, etc.), and organization (i.e., the major league club of that baseball team). The data runs from 1998 through 2019, as 1998 was the first year in which MLB expanded to its current size of 30 organizations, and 2019 was the last full year of data available unaffected by the COVID-19 pandemic.

After scraping this raw data, I eliminated some extraneous columns from each file, as data from minor league teams did not match the format of major league teams, and then combined them into two long files, one for batting data and another for pitching. These files contain statistics describing playing time, player performance, and team played for, with a unique row for each player at each level for each team in each year. This means that if somebody plays at multiple levels in the same year, or changes organizations, they will have multiple entries in a given year.

## Data

*Disclaimer - this is less formal than it will appear in the final paper*

```{r bat_player_summary, echo=FALSE, message=FALSE}

# Compiling some summary statistics

# First getting a table w/ all years w/ >= 20 PA

bat_min_pa_20 <- bat %>% 
  group_by(Name, Year) %>% 
  mutate(pa = sum(PA)) %>% 
  filter(pa >= 20) %>% 
  ungroup()

# Number of players, both with minimums and without

n_bat <- bat %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

n_bat_min <- bat_min_pa_20 %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

# Number of years played, both with minimums and without

n_years <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

n_years_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

# Number of teams played for, both with minimums and without

n_teams <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

n_teams_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

# Number of levels played at, both with minimums and without

n_levels <- bat %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

n_levels_min <- bat_min_pa_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

# Number of plate appearances per year, both with minimums and without

n_pa <- bat %>% 
  group_by(Name, Year) %>% 
  summarize(pa = sum(PA)) %>% 
  select(pa) %>% 
  pull()

n_pa_min <- bat_min_pa_20 %>% 
  group_by(Name, Year) %>% 
  summarize(pa = sum(PA)) %>% 
  select(pa) %>% 
  pull()

bat_player = tibble(
  Measure = c("Years Played", "Organizations", "Levels", "PA per Year"),
  M_1  = c(mean(n_years), mean(n_teams), mean(n_levels), mean(n_pa)),
  SD_1 = c(sd(n_years), sd(n_teams), sd(n_levels), sd(n_pa)),
  M_2  = c(mean(n_years_min), mean(n_teams_min), mean(n_levels_min), mean(n_pa_min)),
  SD_2 = c(sd(n_years_min), sd(n_teams_min), sd(n_levels_min), sd(n_pa_min))
)

kable(bat_player,
      col.names = c("Measure", "M", "SD", "M", "SD"),
      align = c("l", "c", "c", "c", "c"),
      digits = 2,
      caption = "Means and Standard Deviations of Batter Statistics ($n=23,210$, $n_{min}=19,367$)",
      booktabs = T) %>%
  add_header_above(header = c(" " = 1,
                              "No PA Minimum" = 2,
                              "PA >= 20" = 2)) %>% 
  footnote(general = "for PA minimum, player season only included if PA >= 20.",
           general_title = "Note -",
           footnote_as_chunk = TRUE) %>% 
  kable_styling(latex_options = c("striped", "hold_position"))

```

Above are the summary statistics for general player information, including the number of years played, organizations played for, and plate appearances per year. There are two different sets of statistics for each of these: one for the raw data, and one including only those player seasons where the hitter had at least twenty plate appearances, in order to exclude years with injuries or instances where pitchers had plate appearances. We see that when we exclude those seasons, we have about 4,000 fewer players in the data.

From this table, we see that there is a wide range in key statistics across the player pool. Notably, the standard deviation for years played is nearly as large as the mean, indicating here that the distribution of years played is right-skewed, as would be expected in a population where many players have short careers but a select few play at the major league level for an extended period of time.

```{r pitch_player_summary, echo=FALSE, message=FALSE}

# Compiling some summary statistics

# First getting a table w/ all years w/ >= 20 BF

pitch_min_bf_20 <- pitch %>% 
  group_by(Name, Year) %>% 
  mutate(bf = sum(BF)) %>% 
  filter(bf >= 20) %>% 
  ungroup()

# Number of players, both with minimums and without

n_pitch <- pitch %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

n_pitch_min <- pitch_min_bf_20 %>% 
  select(Name) %>% 
  pull() %>% 
  unique() %>% 
  length()

# Number of years played, both with minimums and without

n_years <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

n_years_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Year)) %>% 
  select(count) %>% 
  pull()

# Number of teams played for, both with minimums and without

n_teams <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

n_teams_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Org)) %>% 
  select(count) %>% 
  pull()

# Number of levels played at, both with minimums and without

n_levels <- pitch %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

n_levels_min <- pitch_min_bf_20 %>% 
  group_by(Name) %>% 
  summarize(count = n_distinct(Level)) %>% 
  select(count) %>% 
  pull()

# Number of batters faced per year, both with minimums and without

n_bf <- pitch %>% 
  group_by(Name, Year) %>% 
  summarize(bf = sum(BF)) %>% 
  select(bf) %>% 
  pull()

n_bf_min <- pitch_min_bf_20 %>% 
  group_by(Name, Year) %>% 
  summarize(bf = sum(BF)) %>% 
  select(bf) %>% 
  pull()

pitch_player = tibble(
  Measure = c("Years Played", "Organizations", "Levels", "BF per Year"),
  M_1  = c(mean(n_years), mean(n_teams), mean(n_levels), mean(n_bf, na.rm = TRUE)),
  SD_1 = c(sd(n_years), sd(n_teams), sd(n_levels), sd(n_bf, na.rm = TRUE)),
  M_2  = c(mean(n_years_min), mean(n_teams_min), mean(n_levels_min), mean(n_bf_min)),
  SD_2 = c(sd(n_years_min), sd(n_teams_min), sd(n_levels_min), sd(n_bf_min))
)

kable(pitch_player,
      col.names = c("Measure", "M", "SD", "M", "SD"),
      align = c("l", "c", "c", "c", "c"),
      digits = 2,
      caption = "Means and Standard Deviations of Pitcher Statistics ($n=22,219$, $n_{min}=19,677$)",
      booktabs = T) %>%
  add_header_above(header = c(" " = 1,
                              "No BF Minimum" = 2,
                              "BF >= 20" = 2)) %>% 
  footnote(general = "for BF minimum, player season only included if BF >= 20.",
           general_title = "Note -",
           footnote_as_chunk = TRUE) %>% 
  kable_styling(latex_options = c("striped", "hold_position"))

```

The above table replicates what was done with the batting summary statistics, but for pitchers. The trends here are remarkably similar to those seen in the table for hitters, in both total number of players and their individual metrics. One small difference is that pitchers on average play at slightly more levels than hitters, but this is not an extreme.